---
title: "Customer Segmentation Analysis"
author: "Sanath R"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(ggplot2)
library(dplyr)
library(cluster)
```

## Customer Segmentation Analysis

This project demonstrates how to segment customers based on their purchasing behavior using K-means clustering.

### Data Preparation

```{r data-prep}
# Generate sample customer data
set.seed(456)
n <- 200
customer_data <- data.frame(
  CustomerID = 1:n,
  AnnualIncome = rnorm(n, mean = 60000, sd = 15000),
  SpendingScore = rnorm(n, mean = 50, sd = 25),
  Age = rnorm(n, mean = 35, sd = 12),
  LoyaltyYears = runif(n, 0, 10)
)

# Scale the data for clustering
customer_scaled <- scale(customer_data[, c("AnnualIncome", "SpendingScore", "Age", "LoyaltyYears")])
```

### Determining Optimal Number of Clusters

```{r optimal-clusters, fig.width=8, fig.height=5}
# Calculate WSS for different k values
wss <- sapply(1:10, function(k) {
  kmeans(customer_scaled, centers = k, nstart = 25)$tot.withinss
})

# Plot elbow curve
ggplot(data.frame(k = 1:10, wss = wss), aes(x = k, y = wss)) +
  geom_line() +
  geom_point() +
  theme_minimal() +
  labs(
    title = "Elbow Method for Optimal k",
    x = "Number of Clusters (k)",
    y = "Total Within Sum of Squares"
  )
```

Based on the elbow curve, we'll use k=4 for our clustering.

### K-means Clustering

```{r kmeans, fig.width=9, fig.height=6}
# Perform k-means clustering
k <- 4
km_result <- kmeans(customer_scaled, centers = k, nstart = 25)

# Add cluster assignment to original data
customer_data$Cluster <- as.factor(km_result$cluster)

# Visualize clusters
ggplot(customer_data, aes(x = AnnualIncome, y = SpendingScore, color = Cluster)) +
  geom_point(size = 3, alpha = 0.7) +
  theme_minimal() +
  labs(
    title = "Customer Segments",
    x = "Annual Income",
    y = "Spending Score",
    color = "Cluster"
  ) +
  scale_color_brewer(palette = "Set1")
```

### Customer Segment Profiles

```{r segment-profiles}
# Calculate segment profiles
segment_profiles <- customer_data %>%
  group_by(Cluster) %>%
  summarise(
    Count = n(),
    AvgIncome = mean(AnnualIncome),
    AvgSpending = mean(SpendingScore),
    AvgAge = mean(Age),
    AvgLoyalty = mean(LoyaltyYears)
  )

# Display segment profiles
knitr::kable(segment_profiles, digits = 2)
```

### Marketing Recommendations

Based on the customer segmentation analysis, I recommend the following marketing strategies:

1. **Cluster 1: High Income, High Spenders**
   - Premium product offerings
   - Exclusive loyalty programs
   - Personalized concierge service

2. **Cluster 2: High Income, Low Spenders**
   - Targeted promotions to increase spending
   - Value proposition messaging
   - Product education campaigns

3. **Cluster 3: Low Income, High Spenders**
   - Flexible payment options
   - Mid-tier product recommendations
   - Referral incentives

4. **Cluster 4: Low Income, Low Spenders**
   - Entry-level product focus
   - Special discounts and offers
   - Community-building initiatives

### Conclusion

This segmentation approach allows for more targeted marketing strategies, potentially increasing customer lifetime value and improving marketing ROI. 