---
title: "Sales Analysis Project"
author: "Your Name"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(ggplot2)
library(dplyr)
```

## Sales Analysis Project

This project demonstrates analysis of sales data across different regions and product categories.

### Overview

In this project, I analyzed sales data to identify trends and patterns that could help improve business performance.

### Data Visualization

```{r sales-viz, fig.width=8, fig.height=5}
# Sample data creation
set.seed(123)
regions <- c("North", "South", "East", "West")
categories <- c("Electronics", "Clothing", "Food", "Home Goods")

sales_data <- expand.grid(
  Region = regions,
  Category = categories,
  Year = 2018:2022
)

sales_data$Sales <- runif(nrow(sales_data), 10000, 100000)

# Create a bar chart
ggplot(sales_data, aes(x = Year, y = Sales, fill = Region)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_wrap(~ Category) +
  theme_minimal() +
  labs(
    title = "Sales by Region and Category",
    x = "Year",
    y = "Sales ($)",
    fill = "Region"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

### Key Findings

1. The West region consistently outperforms other regions in Electronics sales
2. Food category shows the most stable growth across all regions
3. Clothing sales experienced a significant drop in 2020 but recovered strongly in 2021

### Recommendations

Based on the analysis, I recommend:

1. Increasing marketing efforts in the South region for Electronics
2. Exploring the success factors in the West region to apply to other regions
3. Developing a more resilient supply chain for the Clothing category

### Conclusion

This analysis provides actionable insights that can help optimize sales strategies across different regions and product categories. 