---
title: "Interactive Portfolio Web App Documentation"
author: "Your Name"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    theme: cosmo
    highlight: tango
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# 1. Introduction

This documentation provides a comprehensive guide to the Interactive Portfolio Web App developed using R Shiny and R Markdown. The application serves as a professional portfolio website showcasing data science projects, resume information, and contact details.

## 1.1 Project Overview

The Interactive Portfolio Web App is designed to be a complete, deployment-ready personal portfolio solution for data scientists and R developers. It features a responsive design, interactive data visualizations, and dynamic content loading capabilities.

## 1.2 Purpose and Scope

The purpose of this application is to:

- Showcase professional skills and experience in an interactive format
- Display data science projects with live visualizations
- Present resume information in a structured, searchable format
- Provide contact information and a functional contact form
- Demonstrate proficiency in R Shiny application development

# 2. System Requirements

## 2.1 Hardware Requirements

- Computer with internet connection
- Minimum 4GB RAM (8GB recommended)
- Sufficient disk space for R and required packages

## 2.2 Software Requirements

- R version 4.0.0 or higher
- RStudio (recommended for development)
- Pandoc version 1.12.3 or higher
- Web browser (Chrome, Firefox, Edge, or Safari)

## 2.3 R Package Dependencies

The following R packages are required:

```{r, eval=FALSE}
# Core functionality
install.packages(c("shiny", "shinythemes", "markdown", "DT", "bslib", "rmarkdown"))

# Enhanced UI and visualization
install.packages(c("plotly", "shinyWidgets", "shinycssloaders", "shinyjs"))

# Data manipulation and visualization
install.packages(c("dplyr", "ggplot2", "viridis", "scales"))
```

# 3. Setting Up the R Environment

## 3.1 Installing R and RStudio

1. Download and install R from [CRAN](https://cran.r-project.org/)
2. Download and install RStudio from [Posit](https://posit.co/download/rstudio-desktop/)
3. Install Pandoc from [pandoc.org](https://pandoc.org/installing.html) if not already included with RStudio

## 3.2 Installing Required Packages

Run the following command in R console to install all required packages:

```{r, eval=FALSE}
install.packages(c("shiny", "shinythemes", "markdown", "DT", "bslib", "rmarkdown",
                  "plotly", "shinyWidgets", "shinycssloaders", "shinyjs",
                  "dplyr", "ggplot2", "viridis", "scales"))
```

## 3.3 Project Structure Setup

The project follows this directory structure:

```
project-portfolio/
├── app.R                 # Main Shiny application
├── /projects             # R Markdown project files
│   ├── Sales Analysis.Rmd
│   └── Customer Segmentation.Rmd
├── /data                 # Data files
│   └── resume.csv        # Resume data in CSV format
├── /www                  # Static assets
│   └── profile.jpg       # Profile picture
└── README.md             # Project documentation
```

# 4. Application Architecture

## 4.1 Core Components

The application is built using the following core components:

1. **UI Layer**: Built with Shiny's `navbarPage` layout with custom CSS styling
2. **Server Logic**: Handles dynamic content loading and user interactions
3. **R Markdown Integration**: Renders project files on demand
4. **Data Tables**: Displays resume information in an interactive format

## 4.2 File Structure and Organization

- **app.R**: Contains both UI and server logic for the Shiny application
- **projects/**: Directory containing R Markdown files that showcase data science projects
- **data/**: Directory containing data files used by the application
- **www/**: Directory containing static assets like images

## 4.3 Data Flow

1. User selects a tab or project from the navigation bar
2. Server processes the request and loads appropriate content
3. For projects, R Markdown files are rendered on-demand using Pandoc
4. For resume information, data is loaded from CSV and displayed in an interactive table

# 5. Implementation Details

## 5.1 Navigation and Layout

The application uses a responsive navigation bar with four main sections:

- **Home**: Introduction and personal information
- **Projects**: Interactive R Markdown project showcase
- **Resume**: Professional experience in tabular format
- **Contact**: Contact information and form

## 5.2 Home Page

The Home page includes:

- Welcome header with profile image
- Brief professional bio
- Skills and expertise section with visual indicators

## 5.3 Projects Page

The Projects page features:

- Dynamic project selector that reads from the projects/ directory
- On-demand rendering of R Markdown files
- Interactive data visualizations within project content
- Category filtering options

## 5.4 Resume Page

The Resume page includes:

- Interactive data table with professional experience
- Filtering and sorting capabilities
- Education history with formatted display
- Certifications and skills section

## 5.5 Contact Page

The Contact page provides:

- Social media and contact links
- Interactive contact form with validation
- Form submission confirmation

# 6. Technologies Used

## 6.1 R Shiny

Shiny is used as the primary framework for creating the interactive web application. Key Shiny features utilized include:

- Reactive programming for dynamic content
- UI components for layout and navigation
- Server-side processing for data handling

## 6.2 R Markdown

R Markdown is used to create the project showcase pages, allowing for:

- Embedded R code execution
- Dynamic data visualization
- Formatted text and documentation
- Conversion to HTML via Pandoc

## 6.3 Additional Libraries

- **bslib**: For enhanced Bootstrap theming
- **DT**: For interactive data tables
- **plotly**: For interactive visualizations
- **shinyjs**: For JavaScript functionality
- **shinycssloaders**: For loading indicators
- **shinyWidgets**: For enhanced UI components

## 6.4 CSS and Styling

Custom CSS is used to enhance the appearance of the application, including:

- Custom fonts from Google Fonts
- Responsive layout adjustments
- Animation and hover effects
- Consistent color scheme and spacing

# 7. Results and Demonstration

## 7.1 Application Features

The completed portfolio application successfully implements all required features:

- Responsive navigation with four main sections
- Dynamic project loading from R Markdown files
- Interactive resume display with filtering and sorting
- Contact form with validation and feedback
- Professional styling with custom CSS

## 7.2 Performance Considerations

The application is optimized for performance in several ways:

- On-demand loading of project content
- Efficient data handling with dplyr
- Loading indicators for long-running operations
- Responsive design for various screen sizes

## 7.3 Browser Compatibility

The application has been tested and is compatible with:

- Google Chrome (recommended)
- Mozilla Firefox
- Microsoft Edge
- Safari

# 8. Deployment

## 8.1 Local Deployment

To run the application locally:

```{r, eval=FALSE}
# Set working directory to the project folder
setwd("path/to/project-portfolio")

# Run the application
shiny::runApp()
```

## 8.2 Deployment to shinyapps.io

To deploy the application to shinyapps.io:

1. Install the rsconnect package:

```{r, eval=FALSE}
install.packages("rsconnect")
```

2. Set up your shinyapps.io account credentials:

```{r, eval=FALSE}
rsconnect::setAccountInfo(
  name = "your-account-name",
  token = "YOUR_TOKEN",
  secret = "YOUR_SECRET"
)
```

3. Deploy the application:

```{r, eval=FALSE}
rsconnect::deployApp(
  appName = "portfolio-app",
  appTitle = "My Portfolio"
)
```

# 9. Customization Guide

## 9.1 Personal Information

To customize the personal information:

1. Edit the text content in the Home tab section of app.R
2. Replace the profile.jpg file in the www/ directory
3. Update social media links in the Contact tab section

## 9.2 Adding Projects

To add new projects:

1. Create a new R Markdown file in the projects/ directory
2. Follow the format of existing project files
3. The application will automatically detect and display the new project

## 9.3 Updating Resume

To update resume information:

1. Edit the resume.csv file in the data/ directory
2. Follow the existing format with columns for Position, Company, Location, etc.
3. The changes will be reflected automatically when the application runs

## 9.4 Styling Customization

To customize the application's appearance:

1. Modify the CSS styles in the tags$style section of app.R
2. Change the theme by modifying the bs_theme parameters
3. Update color schemes and fonts as desired

# 10. Conclusion and Learning Outcomes

## 10.1 Project Summary

This Interactive Portfolio Web App demonstrates the power of R Shiny for creating professional, interactive web applications. The combination of Shiny, R Markdown, and modern web technologies results in a polished, functional portfolio that effectively showcases data science skills and projects.

## 10.2 Learning Outcomes

Through the development of this application, several key skills were demonstrated:

- Building interactive web applications with R Shiny
- Integrating R Markdown documents into Shiny applications
- Creating responsive layouts with custom CSS
- Implementing interactive data visualizations
- Designing user-friendly interfaces with modern UI components
- Optimizing application performance and user experience

## 10.3 Future Enhancements

Potential future enhancements for the application include:

- Integration with GitHub API to automatically pull project information
- Adding user authentication for admin features
- Implementing a blog section with regular updates
- Adding analytics to track visitor interactions
- Creating a dark mode theme option

# 11. References

- [Shiny Documentation](https://shiny.posit.co/r/reference/shiny/latest/)
- [R Markdown Documentation](https://rmarkdown.rstudio.com/)
- [bslib Documentation](https://rstudio.github.io/bslib/)
- [DT Package Documentation](https://rstudio.github.io/DT/)
- [shinyapps.io Deployment Guide](https://docs.posit.co/shinyapps.io/) 